\documentclass{manual}
\usepackage{palatino}
\renewcommand{\ttdefault}{cmtt}
\renewcommand{\sfdefault}{cmss}
\newcommand{\myhdl}{\protect \mbox{MyHDL}}
\usepackage{graphicx}

\title{The \myhdl\ manual}

\input{boilerplate}

\makeindex

\begin{document}

\maketitle

\input{copyright}

\begin{abstract}

\noindent

The goal of the \myhdl{} project is to empower hardware designers with
the elegance and simplicity of the Python language.

\myhdl{} is a free, open-source (LGPL) package for using Python as a
hardware description and verification language. Python is a very high
level language, and hardware designers can use its full power to model
and simulate their designs. Moreover, \myhdl{} can convert a design to
Verilog. In combination with an external synthesis tool, it provides a
complete path from Python to a silicon implementation.

\emph{Modeling}


Python's power and clarity make \myhdl{} an ideal solution for high level
modeling. Python is famous for enabling elegant solutions to complex
modeling problems. Moreover, Python is outstanding for rapid
application development and experimentation.

The key idea behind \myhdl{} is the use of Python generators to model
hardware concurrency. Generators are best described as resumable
functions. In \myhdl{}, generators are used in a specific way so that
they become similar to always blocks in Verilog or processes in VHDL.

A hardware module is modeled as a function that returns any number of
generators. This approach makes it straightforward to support features
such as arbitrary hierarchy, named port association, arrays of
instances, and conditional instantiation.

Furthermore, \myhdl{} provides classes that implement traditional
hardware description concepts. It provides a signal class to support
communication between generators, a class to support bit oriented
operations, and a class for enumeration types.

\emph{Simulation and Verification}

The built-in simulator runs on top of the Python interpreter. It
supports waveform viewing by tracing signal changes in a VCD file.

With \myhdl{}, the Python unit test framework can be used on hardware
designs. Although unit testing is a popular modern software
verification technique, it is not yet common in the hardware design
world, making it one more area in which \myhdl{} innovates.

\myhdl{} can also be used as hardware verification language for VHDL and
Verilog designs, by co-simulation with traditional HDL simulators.

\emph{Conversion to Verilog}

The converter to Verilog works on an instantiated design that has been
fully elaborated. Consequently, the original design structure can be
arbitrarily complex.

The converter automates certain tasks that are tedious or hard in
Verilog directly. Notable features are the possibility to choose
between various FSM state encodings based on a single attribute, the
mapping of certain high-level objects to RAM and ROM descriptions, and
the automated handling of signed arithmetic issues.



\end{abstract}

\tableofcontents

\input{background.tex}
\input{intro.tex}
\input{modeling.tex}
\input{unittest.tex}
\input{cosimulation.tex}

\chapter{Conversion to Verilog\label{conv}}
\input{conversion.tex}

\input{reference.tex}

\input{MyHDL.ind}

\end{document}
